#STAGE 1 TASK 
 #1.Write a function for translating DNA to protein
# Define the genetic code as a named vector
genetic_code <- c(
  "ATA" = "I", "ATC" = "I", "ATT" = "I", "ATG" = "M",
  "ACA" = "T", "ACC" = "T", "ACG" = "T", "ACT" = "T",
  "AAC" = "N", "AAT" = "N", "AAA" = "K", "AAG" = "K",
  "AGC" = "S", "AGT" = "S", "AGA" = "R", "AGG" = "R",
  "CTA" = "L", "CTC" = "L", "CTG" = "L", "CTT" = "L",
  "CCA" = "P", "CCC" = "P", "CCG" = "P", "CCT" = "P",
  "CAC" = "H", "CAT" = "H", "CAA" = "Q", "CAG" = "Q",
  "CGA" = "R", "CGC" = "R", "CGG" = "R", "CGT" = "R",
  "GTA" = "V", "GTC" = "V", "GTG" = "V", "GTT" = "V",
  "GCA" = "A", "GCC" = "A", "GCG" = "A", "GCT" = "A",
  "GAC" = "D", "GAT" = "D", "GAA" = "E", "GAG" = "E",
  "GGA" = "G", "GGC" = "G", "GGG" = "G", "GGT" = "G",
  "TCA" = "S", "TCC" = "S", "TCG" = "S", "TCT" = "S",
  "TTC" = "F", "TTT" = "F", "TTA" = "L", "TTG" = "L",
  "TAC" = "Y", "TAT" = "Y", "TAA" = "*", "TAG" = "*",
  "TGC" = "C", "TGT" = "C", "TGA" = "*", "TGG" = "W"
)

# Function to translate DNA to protein
translate_dna <- function(dna) {
  # Split the DNA sequence into codons (groups of 3 bases)
  codons <- substring(dna, seq(1, nchar(dna), by = 3), seq(3, nchar(dna), by = 3))

  
  # Translate each codon to an amino acid using the genetic code
  protein <- sapply(codons, function(codon) genetic_code[[codon]])
  
  # Combine the amino acids into a protein sequence
  return(paste(protein, collapse = ""))
}

# Example usage
result <- translate_dna("ATGGGGAAGGTGAAGGTCGGAGTCAACGGATTTGGTCGTATTGGGCGCCTGGTCACCAGGGCTGCTTTTAACTCTGGTGGGGGTCACTCTGGGATCTGACTTGTGGAAGGCTCATGGGGGAGGTGGGCCCTGCTGCTGGTGGTTCTGAGTATGGCACGGACTAATTGGAGAAAGCAGCACGGCTCCTTTCGTGGTCTTCACCCCTCCTGCTGCTGCTGCTGCTAC")
print(result)

```

```{r}
# 2. Write a function that simulates and generates a logistic population growth curve. Your function should include 2 extra parameters that randomize the length of the lag phase and the exponential phase [See population curve here] . Most living populations follow a logistic population growth. Therefore, your growth curve can be: Population Size vs Time, Cell density vs Time, OD vs Time, CFU vs Time, etc. Using your function, generate a dataframe with 100 different growth curves

# Simple logistic growth function (time, 78, 0.4, 7, 6) -> Population
logistic_growth <- function(time, K, r, P0, lag) {
  if (time < lag) {
    return(P0) # Population stays constant during lag phase
  }
  else{
    # Logistic growth equation
    return(K / (1 + ((K - P0) / P0) * exp(-r * (time - lag))))
  }
}

# Function to generate multiple growth curves
generate_growth_curves <- function(n_curves = 100, time_points = 50) {
  all_curves <- data.frame() # Initialize an empty dataframe
  
  for (i in 1:n_curves) {
    # Randomize parameters for each curve
    K <- runif(1, 50, 100)  # Carrying capacity - 78
    r <- runif(1, 0.1, 0.5) # Growth rate - 0.4
    P0 <- runif(1, 1, 10)   # Initial population - 7
    lag <- sample(1:10, 1)  # Lag phase duration - 6
    
    # Generate time points and population sizes 0 - 50
    time <- seq(0, time_points, by = 1)
    population <- sapply(time, function(t) logistic_growth(t, K, r, P0, lag))
    
    # Combine into a dataframe
    curve_df <- data.frame(Curve_ID = i, Time = time, Population = population)
    
    # Append to the main dataframe
    all_curves <- rbind(all_curves, curve_df)
  }
  
  return(all_curves)
}

# Generate 100 growth curves
growth_curves <- generate_growth_curves()

# View the first few rows
head(growth_curves)


```

```{r}
# 3. Write a function for determining the time to reach 80% of the maximum growth; usually the carrying capacity
# Logistic growth function
logistic_growth <- function(t, K, r, P0) {
  return(K / (1 + (K - P0) / P0 * exp(-r * t)))
}

# Function to find time to reach 80% of carrying capacity
time_to_80_percent <- function(K, r, P0) {
  target_population <- 0.8 * K  # 80% of carrying capacity
  
  # Loop through time steps to find the time
  for (t in 0:100) {
    population <- logistic_growth(t, K, r, P0)
    if (population >= target_population) {
      return(t)
    }
  }
  return(NA)  # If not reached within 100 time steps
}

# Example usage
K <- 100     # Carrying capacity
r <- 0.3     # Growth rate
P0 <- 5      # Initial population

time_80 <- time_to_80_percent(K, r, P0)
cat("Time to reach 80% of carrying capacity:", time_80, "\n")

```

```{r}
# 4. Finally, write a function for calculating the hamming distance between your Slack username and twitter/X handle (synthesize if you don’t have one). Feel free to pad it with extra words if they are not of the same length.
# Function to calculate Hamming distance
hamming_distance <- function(str1, str2) {
  # Pad the shorter string with spaces if they are not the same length
  max_length <- max(nchar(str1), nchar(str2))
  str1 <- paste0(str1, rep(" ", max_length - nchar(str1)))
  str2 <- paste0(str2, rep(" ", max_length - nchar(str2)))
  
  # Count the number of differing characters
  diff_count <- sum(strsplit(str1, NULL)[[1]] != strsplit(str2, NULL)[[1]])
  
  return(diff_count)
}

# Example usage
slack_username <- "aspiara2468102"
x_username <- "aspiararoosmin"
distance <- hamming_distance(slack_username, x_username)
print(distance)

```


